---
title: "leaflet"
author: "Maggie Klope"
date: "1/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(leaflet)
library(leaflet.minicharts)
library(sf)

```

## Leaflet: Easily create interactive maps

### Combatible with a lot of different data sources

- lat/long matrices
- data frame with lat/long columns
- data types from the sp package (SpatialPoints, lines, Polygons, SpatialPolygons)

## Basic Map with one point
```{r}
ca_capitol <- leaflet() %>%
  addTiles() %>% #loads default map
  addMarkers(lat = 38.576572, lng = -121.4930107, popup = "CA State Capitol")

ca_capitol

```

## Using shapefiles
```{r}
# use sf to read the shapefile
birds <- st_read("ds78.shp") %>% 
  filter(LONGITUDE != 0.0000) %>%  # removing a few data points that were missing lat/long
  filter(REGION == "Southwest")

bird_map <- leaflet(data = birds) %>% 
  addTiles() %>% 
  addMarkers(lat = birds$LATITUDE, lng = birds$LONGITUDE)


```

## Basic Map Customization

### Lots of different backgrounds maps you can use with addProviderTiles()

add link here

```{r}
# from more informative
bird_map %>% addProviderTiles(providers$Esri.WorldImagery)

# to more artistic
bird_map %>% addProviderTiles(providers$Stamen.Watercolor)

# to more... metal
bird_map %>% addProviderTiles(providers$Thunderforest.SpinalMap) # not working


```

## cluster maps
```{r}
leaflet(data = birds) %>%
  addTiles() %>%
  addMarkers(lat = birds$LATITUDE, lng = birds$LONGITUDE,
  clusterOptions = markerClusterOptions()
)

```

## can do fun things like add custom icons
```{r}

pelicans <- birds %>% 
  filter(BRPN == "Y")

pelican_icon <- makeIcon(iconUrl = "pelican.png", iconHeight = 20, iconWidth = 20)

leaflet(data = pelicans) %>% 
  addTiles() %>% 
  addMarkers(lat = pelicans$LATITUDE, lng = pelicans$LONGITUDE, icon = pelican_icon)
  
```

## Using the leaflet.minicharts package, you can add figures over your map

```{r}
bird_matrix <- as.tibble(birds)

leaflet(data = bird_matrix) %>% 
  addTiles() %>% 
  addMinicharts(
    lng = bird_matrix$LONGITUDE,
    lat = bird_matrix$LATITUDE,
    type = "bar",
    chartdata = bird_matrix[, c("SCORE", "THREAT")])

leaflet(data = bird_matrix) %>% 
  addTiles() %>% 
  addMinicharts(
    lng = bird_matrix$LONGITUDE,
    lat = bird_matrix$LATITUDE,
    type = "pie",
    chartdata = bird_matrix[, c("SCORE", "THREAT")])

leaflet(data = bird_matrix) %>% 
  addTiles() %>% 
  addMinicharts(
    lng = bird_matrix$LONGITUDE,
    lat = bird_matrix$LATITUDE,
    showLabels = TRUE,
    chartdata = bird_matrix$SCORE)

```

## Compatible with Shiny


```{r


```


