---
title: "leaftlet_runthrough"
author: "Maggie Klope"
date: "2/9/2022"
output: html_document
---

## Leaflet: Easily create interactive maps

- Uses the [Leaflet](https://leafletjs.com/index.html) open-source JavaScript library
- Quick way to make interactive maps
- Can be embedded into R Markdown and Shiny apps
- Compatible with a lot of data types (data frames with lat/long columns, sp and sf package spatial objects)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# install packages if you have not already
# install.packages("tidyverse")
# install.packages("leaflet")
# install.packages("leaflet.minicharts")
# install.packages("sf")

library(tidyverse)
library(leaflet)
library(leaflet.minicharts)
library(sf)

```

### Basic Map with one point
```{r}

# Using the lat/long for the Methuselah Tree (oldest non-clonal tree in the world!)
methuselah <- 
  
# adding more features

# lots of different privider tiles: https://leaflet-extras.github.io/leaflet-providers/preview/
methuselah %>% addProviderTiles(providers$Esri.WorldImagery)

methuselah %>% addProviderTiles(providers$Stamen.Watercolor)
  
```

### Using shapefile

- Audubon Society occurrence surveys
- Includes a count of sensitive and threatened species
- Assigns a Score value for the taxa that exceed the population threshold of 1% of global or 10% of state population

```{r}

# use sf to read the shapefile
birds <- 

# creating a point map
bird_map <- 

```

### Creating a cluster map
```{r}


```

### Creating custom icons
```{r}

# making a dataset for locations where brown pelicans (BRPN) were detected
pelicans <- 

# using the makeIcon() function to load a .png file
pelican_icon <- 

# creating map with custom icon


```

### Using the leaflet.minicharts package, you can add figures over your map	
```{r}	
bird_matrix <- as.tibble(birds)

# adding barcharts	
leaflet(data = bird_matrix) %>% 	
  addTiles() %>% 	
  addMinicharts(	
    lng = bird_matrix$LONGITUDE,	
    lat = bird_matrix$LATITUDE,	
    chartdata = bird_matrix[, c("SCORE", "THREAT")],# just using the Score and Threat columns	
    type = "bar") 

# adding piecharts	


# adding label	

```	

### Maps with Polylines	

```{r}	
# loading example dataset from the leaflet package	
storm_data <- atlStorms2005	

# Create a continuous palette function	
palette <- 

# Creating a map
storm <- 

storm	

# some extra features	
storm %>% 	
  addGraticule() %>% 	
  addTerminator()	
```	

### Compatible with Shiny	

- See the leaflet_shiny_app.R file